---
title: "Data & Methods"
---

## Data Sources

All data were extracted from official ANVUR documents published on 28 May 2026:

| Document | Content | Reference |
|----------|---------|-----------|
| *Rapporto Istituzione — Università degli Studi di Padova* | Institutional report with department-level indicators (Tables 2.1–2.4) | [ANVUR VQR 2020–2024 Results](https://www.anvur.it/attivita/vqr/vqr-2020-2024/) |
| *Rapporto Finale ANVUR VQR 2020–2024* | National-level methodology and indicator definitions | [ANVUR Final Report](https://www.anvur.it/attivita/vqr/vqr-2020-2024/) |

### What the ANVUR tables contain

The Padova institutional report includes the following tables, which form the basis of this analysis:

- **Table 2.1 (R1)**: Quality indicator for tenured staff, by department and disciplinary area (*area CUN*).
- **Table 2.2 (R2)**: Quality indicator for newly recruited / promoted staff.
- **Table 2.3 (R1_2)**: Quality indicator for all evaluated staff combined.
- **Table 2.4**: Summary indicators (H, R, IRD) aggregated across disciplinary areas for all 32 departments.

## Data Extraction

Data were extracted in two ways:

1. **Area-level indicators** (R1, R2, R1_2 by department within Areas 05 and 06): Manually transcribed from the PDF report into [`vqr_area_data.csv`](https://github.com/Russlewisbo/VQR_2020-2024/blob/main/vqr_area_data.csv). The pipeline script also includes a PDF parsing function (`parse_area_tables()`) using `pdftools` and regex, though the verified CSV is used by default.

2. **Department-level indicators** (all 32 departments): Transcribed into [`vqr_ird_data.csv`](https://github.com/Russlewisbo/VQR_2020-2024/blob/main/vqr_ird_data.csv) from Table 2.4 of the institutional report.

::: {.callout-warning}
## Data extraction caveats

- **R1 data for Area 05** includes only 5 of the 7 departments. BCA and DNS are excluded because they had fewer than 10 expected products from tenured staff — an ANVUR reporting threshold, not a data gap.
- All R1 values were initially missing from the CSV and were manually added from pages 11–12 of the ANVUR PDF during the analysis. The values should be cross-checked against the original source if exact reproducibility is critical.
:::

## Indicator Definitions

### Quality indicator R

The R indicator for a department in a given area is:

$$
R = \frac{\bar{v}_{\text{dept}}}{\bar{v}_{\text{national}}}
$$

where $$\bar{v}$$ is the average evaluation score of submitted products. R > 1 means the department's average quality exceeds the national average for that area.

### Size-weight indicator H

$$
H = \frac{n_{\text{dept}}}{N_{\text{institution}}}
$$

where $$n_{\text{dept}}$$ is the number of expected products from the department and $$N_{\text{institution}}$$ is the total across the institution. H reflects the department's relative contribution to the institution's evaluated output.

### Composite indicator IRD

$$
IRD = H \times R
$$

This product means that IRD scales with both quality (R) and size (H). A small department with high R can still have a low IRD, because H is small. This is a deliberate design choice by ANVUR: IRD is intended to capture *total research impact*, not quality per capita.

::: {.callout-important}
## Critical note on IRD interpretation

IRD is sometimes used in internal benchmarking as though it measures departmental "performance." This is misleading. A department with R = 1.26 (top quality) but only 157 expected products will always score lower on IRD than a department with R = 1.05 but 350 products. **IRD reflects departmental scale as much as — or more than — research quality.** This is explored in detail in the [Results](results.qmd#sec-ird-scatter).
:::

## Reproducibility

The full analysis pipeline is contained in [`vqr_pipeline.R`](https://github.com/Russlewisbo/VQR_2020-2024/blob/main/vqr_pipeline.R). To reproduce all figures:
```r
source("vqr_pipeline.R")
```

The figures on the Results page are generated directly from the embedded R code in this Quarto document, using the same data files and plotting logic.

### Software environment

```{r}
#| label: session-info
#| code-fold: true
#| code-summary: "Session info"
sessionInfo()
```

### Required packages

```{r}
#| eval: false
pkgs <- c("dplyr", "tidyr", "ggplot2", "patchwork",
          "forcats", "scales", "ggrepel", "readr")
install.packages(pkgs)
```
